# A script for manipulating gene expression data 
#setwd("E:/Study/Projects/BIOINFORMATICs/Gene expression data manuplation _1")

# loda libraries 

```{r}
library(dplyr)
library(tidyverse)
library(GEOquery)
```


#read the data 
```{r}
data <- read.csv("GSE183947_fpkm.csv")
```

#To check the dimension of data 

```{r}
dim(data)
```
#get matadata or download 

```{r}
GSE <- getGEO(GEO = 'GSE183947', GSEMatrix = TRUE)
```
#To see about the data 

```{r}
GSE
```


```{r}
metadata <- pData(phenoData(GSE[[1]]))
head(metadata)

```

# To extract the important columns

#```{r}
#metadat.subset <- select(metadata,c(1,10,11,17))
#```

#      or  by using pipe(%>% use 4 going inside ) function
```{r}

metadat.modified <- metadata %>%
  select(1,10,11,17)%>%
  rename(tissue = characteristics_ch1)%>% # rename(function) is use to chane the name of column 
  rename(metastasis = characteristics_ch1.1)%>%
  mutate(tissue = gsub("tissue:","",tissue))%>% # mutate (function is used to change or remove in column)
  mutate(metastasis = gsub("metastasis:","", metastasis)) # gsub means get subsitute 



```


#reshape the data

```{r}
data.long <- data %>%
  rename(gene = X)%>%
  gather(key = "sample" ,value = "FPKM" ,-gene) # gathe is fuction which convert to wide format to long format 
  
```

```{r}
# Save the data to a CSV file
#write.csv(data.long, file = "E:/Study/Projects/BIOINFORMATICs/Gene expression data manuplation _1.csv", row.names = FALSE)

```


# join dataframes  == data.long + metadata.modified 


```{r}
Data.long <- data.long %>%
  left_join(.,metadat.modified, by = c("sample" = "description")) 
head(data.long)
```



```{r}
# Save the data to a CSV file
write.csv(data.long, file = "E:/Study/Projects/BIOINFORMATICs/Data.long _1.csv", row.names = FALSE)

```


#explore data 



```{r}
Data.long %>%
  filter(gene == "TSPAN6" | gene == "TNMD")%>% #filter(function) is used extract raws
  group_by(gene,tissue)%>%
  summarize(mean_FPKM = mean(FPKM),
            median_FPKM = median((FPKM)))

```
  



